---
title: "Plots"
author: "aaron till"
date: "4/19/2018"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(stringr)
library(plyr)
library(ggthemes)
```





# Data Import

```{r}

setwd("~/counting-the-homeless/Data")

Affordable_Housing_10_14 <- read.csv("HAI_map_2010-14.csv")  %>%
  dplyr::rename('% Housing Units per ELI in 2014' = 'per100') %>%
  dplyr::rename('Housing Units 2014' = 'units')

Affordable_Housing_05_09 <- read.csv("HAI_map_2005-09.csv")  %>%
  dplyr::rename('% Housing Units per ELI in 2009' = 'per100') %>%
  dplyr::rename('Housing Units 2009' = 'units')



#Evictions <- read_xlsx('eviction_lab_export.xlsx') not using because bad lvl of aggregation

#Oregon_Housing_Prices <- read_xlsx('2017-OOR-OR.xlsx') not using because included in census



PIT_State <- read_xlsx("2007-2017-PIT-Counts-by-State.xlsx")

Helper_State_Abrvs <- read.csv("states.csv.csv")

```

# Joining the affordable housing

```{r}
Affordable_Housing <- full_join(Affordable_Housing_05_09, Affordable_Housing_10_14, by = c('countyname' = 'countyname', 'State_Name' = 'State_Name')) %>%
  mutate(countyname2 = str_remove(countyname, pattern = ' County')) %>%
  select(countyname, countyname2, State_Name, `% Housing Units per ELI in 2014`, `% Housing Units per ELI in 2009`, `Housing Units 2014`, `Housing Units 2009` )
```


#PIT COC - Expanding years

```{r}

setwd("~/counting-the-homeless/Data")


PIT_CoC_2017 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 1)
PIT_CoC_2016 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 2)
PIT_CoC_2015 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 3)
PIT_CoC_2014 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 4)
PIT_CoC_2013 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 5)
PIT_CoC_2012 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 6)
PIT_CoC_2011 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 7)
PIT_CoC_2010 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 8)
PIT_CoC_2009 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 9)
PIT_CoC_2008 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 10)
PIT_CoC_2007 <- read_xlsx("2007-2017-PIT-Counts-by-CoC.XLSX.xlsx", sheet = 11)

PIT_CoC <- join_all(list(PIT_CoC_2007, PIT_CoC_2008, PIT_CoC_2009, PIT_CoC_2010, PIT_CoC_2011, PIT_CoC_2012, PIT_CoC_2013, PIT_CoC_2014, PIT_CoC_2015, PIT_CoC_2016, PIT_CoC_2017)) %>%
  mutate(countyname2 = str_remove(`CoC Name`, pattern = ' CoC'))  # INCORPORATED REGIONS NOT COUNTIES
```

PIT_CoC <- PIT_CoC %>%
  dplyr::rename('CoC_Name' = 'CoC Name') %>% #dplyr:: neccesary because plyr analog
  dplyr::rename('Unsheltered_2017' =  'Unsheltered Homeless, 2017' )%>%
  dplyr::rename('Total_Homeless_2017' = 'Homeless Individuals, 2017') %>%
  dplyr::rename('Unsheltered_2016' =  'Unsheltered Homeless, 2016' )%>%
  dplyr::rename('Total_Homeless_2016' = 'Homeless Individuals, 2016') %>%
  dplyr::rename('Unsheltered_2015' =  'Unsheltered Homeless, 2015' )%>%
  dplyr::rename('Total_Homeless_2015' = 'Homeless Individuals, 2015') %>%
  dplyr::rename('Unsheltered_2014' =  'Unsheltered Homeless, 2014' )%>%
  dplyr::rename('Total_Homeless_2014' = 'Homeless Individuals, 2014') %>%
  dplyr::rename('Unsheltered_2013' =  'Unsheltered Homeless, 2013' )%>%
  dplyr::rename('Total_Homeless_2013' = 'Homeless Individuals, 2013') %>%
  dplyr::rename('Unsheltered_2012' =  'Unsheltered Homeless, 2012' )%>%
  dplyr::rename('Total_Homeless_2012' = 'Homeless Individuals, 2012') %>%
  dplyr::rename('Unsheltered_2011' =  'Unsheltered Homeless, 2011' )%>%
  dplyr::rename('Total_Homeless_2011' = 'Homeless Individuals, 2011') %>%
  dplyr::rename('Unsheltered_2010' =  'Unsheltered Homeless, 2010' )%>%
  dplyr::rename('Total_Homeless_2010' = 'Homeless Individuals, 2010') %>%
  dplyr::rename('Unsheltered_2009' =  'Unsheltered Homeless, 2009' )%>%
  dplyr::rename('Total_Homeless_2009' = 'Homeless Individuals, 2009') %>%
  dplyr::rename('Unsheltered_2008' =  'Unsheltered Homeless, 2008' )%>%
  dplyr::rename('Total_Homeless_2008' = 'Homeless Individuals, 2008') %>%
  dplyr::rename('Unsheltered_2007' =  'Unsheltered Homeless, 2007' )%>%
  dplyr::rename('Total_Homeless_2007' = 'Homeless Individuals, 2007') %>%
DECIDEd NOT TO RENAME


# County Tidying of Affordable Housing - Oregon Only

```{r}

setwd("~/counting-the-homeless/Data")

Affordable_Housing_Oregon <- Affordable_Housing %>%
  filter(State_Name == 'Oregon') #%>%
  #filter(!is.na(Total)) dont need this anymore becuase filtered in earlier step

Affordable_Housing_Oregon$Area <- Affordable_Housing_Oregon$countyname2

Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Hood River')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Jefferson')] <- 'Central_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Crook')] <- 'Central_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Tillamook')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Curry')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Wasco')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Columbia')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Malheur')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Union')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Clatsop')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Lincoln')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Umatilla')] <- 'Rural_Oreogn'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Coos')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Polk')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Yamhill')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Klamath')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Josephine')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Douglas')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Linn')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Deschutes')] <- 'Central_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Benton')] <- 'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Jackson')] <- 'Jackson'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Clackamas')] <- 'Clackamas'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Marion')] <-'Rural_Oregon'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Washington')] <- 'Washington'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Lane')] <- 'Lane'
Affordable_Housing_Oregon$Area[which(Affordable_Housing_Oregon$countyname2 == 'Multnomah')] <- 'Multnomah' 


Affordable_Housing_Oregon <- Affordable_Housing_Oregon %>%
  group_by(Area) %>%
  dplyr::summarise(`Housing Units 2014` = sum(`Housing Units 2014`), `% Housing Units per ELI in 2014` = mean(`% Housing Units per ELI in 2014`), `Housing Units 2009` = sum(`Housing Units 2009`), `% Housing Units per ELI in 2009` = mean(`% Housing Units per ELI in 2009`)) # also must specify dplyr

```

#County tidying of PIT - Oregon Only 

```{r}

setwd("~/counting-the-homeless/Data")

PIT_CoC_Oregon<- PIT_CoC %>%
  filter(countyname2 == 'Eugene/Springfield/Lane County'| countyname2 == 'Portland-Gresham-Multnomah County'| countyname2 == 'Medford/Ashland/Jackson County'| countyname2 == 'Central Oregon'| countyname2 == 'Oregon Balance of State'| countyname2 == 'Hillsboro/Beaverton/Washington County'| countyname2 == 'Clackamas County') 


PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Eugene/Springfield/Lane County')] <- 'Lane'
PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Portland-Gresham-Multnomah County')] <- 'Multnomah'
PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Medford/Ashland/Jackson County')] <- 'Jackson'
PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Central Oregon')] <- 'Central_Oregon'
PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Oregon Balance of State')] <- 'Rural_Oregon'
PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Hillsboro/Beaverton/Washington')] <- 'Washington'
PIT_CoC_Oregon$Area[which(PIT_CoC_Oregon$countyname2 =='Clackamas County')] <- 'Clackamas'




PIT_Housing_Oregon_Merge <- merge(Affordable_Housing_Oregon, PIT_CoC_Oregon)



```







# State Tidying
```{r}


Affordable_Housing_State <- Affordable_Housing %>%
  group_by(State_Name) %>%
  dplyr::summarise(`Housing Units 2014` = mean(`Housing Units 2014`, na.rm = TRUE), `% Housing Units per ELI in 2014` = mean(`% Housing Units per ELI in 2014`, na.rm = TRUE), `Housing Units 2009` = mean(`Housing Units 2009`, na.rm = TRUE), `% Housing Units per ELI in 2009` = mean(`% Housing Units per ELI in 2009`, na.rm = TRUE))

Intermediate_Housing_State_Abrvs <-full_join(Affordable_Housing_State, Helper_State_Abrvs, by = c('State_Name' = 'State'))

PIT_Housing_State_Merge <- inner_join(Intermediate_Housing_State_Abrvs, PIT_State, by = c('Abbreviation' = 'state')) %>%
  dplyr::rename('Change_2007' = 'Change in Total Homelessness, 2007-2017') %>%
  dplyr::rename('Change_2008' = 'Change in Total Homelessness, 2008-2017') %>%
  dplyr::rename('Change_2009' = 'Change in Total Homelessness, 2009-2017') %>%
  dplyr::rename('Change_2010' = 'Change in Total Homelessness, 2010-2017') %>%
  dplyr::rename('Change_2011' = 'Change in Total Homelessness, 2011-2017') %>%
  dplyr::rename('Change_2012' = 'Change in Total Homelessness, 2012-2017') %>%
  dplyr::rename('Change_2013' = 'Change in Total Homelessness, 2013-2017') %>%
  dplyr::rename('Change_2014' = 'Change in Total Homelessness, 2014-2017') %>%
  dplyr::rename('Change_2015' = 'Change in Total Homelessness, 2015-2017') %>%
  dplyr::rename('Change_2016' = 'Change in Total Homelessness, 2016-2017') 



```





#Census Data Pt.1 
```{r eval=FALSE}
setwd("~/counting-the-homeless/Data")

OR_county_census_filtered <- read.csv("R11689837_SL050.csv")
OR_county_census_filtered %>%
  select(1:4, 56:107) %>%
  dplyr::rename('Total Population' = 'SE_T001_001') %>%
  dplyr::rename('Civilian Population 16 to 19 Years' = 'SE_T030_001') %>%
  dplyr::rename('Not High School Graduate, Not Enrolled (Dropped Out)' = 'SE_T030_002') %>%
  dplyr::rename('High School Graduate, or Enrolled (in School)' = 'SE_T030_003') %>%
  dplyr::rename('Population 16 Years and Over' = 'SE_T033_001') %>% 
  dplyr::rename('In Labor Force' = 'SE_T033_002') %>%
  dplyr::rename('In Armed Forces' = ' SE_T033_003') %>%
  dplyr::rename('Civilian' = 'SE_T033_004') %>%
  dplyr::rename('Employed' = 'SE_T033_005') %>%
  dplyr::rename('Unemployed' = 'SE_T033_006') %>%
  dplyr::rename('Not in Labor Force' = 'SE_T033_007') %>%
  ###Unemployment data(T37) left out since it is included in T33
   dplyr::rename('Median Household Income (In 2016 Inflation Adjusted Dollars)' = 'SE_T057_001') %>%
   dplyr::rename('Per Capita Income (In 2016 Inflation Adjusted Dollars)' = 'SE_T083_001') %>%
   dplyr::rename('Gini Index' = 'SE_T157_00') %>%
   dplyr::rename('Housing Units' = 'SE_T093_001') %>%
   dplyr::rename('Occupied Housing Units' = 'SE_T099_001') %>%
  ###unsure if rest of T99(housing fuel stuff) is useful 
  dplyr::rename('Median Gross Rent' = 'SE_T099_001') %>%
  dplyr::rename('Average Gross Rent for Renter-Occupied Housing Units' = 'SE_T099_001') %>%
  ###avg and median are actually somewhat different
  dplyr::rename('Occupied Housing Units' = 'SE_T099_001') %>%
  dplyr::rename('Occupied Housing Units' = 'SE_T099_001') %>%
  dplyr::rename('Occupied Housing Units' = 'SE_T099_001') %>%
  dplyr::rename('Occupied Housing Units' = 'SE_T099_001') 
```

#OR county Census Data we'll actually probably use
```{r}
setwd("~/counting-the-homeless/Data")

perc_or_census <- read.csv("R11689837_SL050 (1).csv")
perc_or_census_cleaned <- perc_or_census %>%
  select(1:4, 56:143) %>%
  dplyr::rename('Total Population' = 'SE_T001_001') %>%
  dplyr::rename('Civilian Population 16 to 19 Years' = 'SE_T030_001') %>%
  dplyr::rename('% Not High School Graduate, Not Enrolled (Dropped Out)' = 'PCT_SE_T030_002') %>%
  dplyr::rename('% High School Graduate, or Enrolled (in School)' = 'PCT_SE_T030_003') %>%
  dplyr::rename('Population 16 Years and Over' = 'SE_T033_001') %>% 
  dplyr::rename('% In Labor Force' = 'PCT_SE_T033_002') %>%
  dplyr::rename('% In Armed Forces' = 'PCT_SE_T033_003') %>%
  dplyr::rename('% Civilian' = 'PCT_SE_T033_004') %>%
  dplyr::rename('% Employed' = 'PCT_SE_T033_005') %>%
  dplyr::rename('% Unemployed' = 'PCT_SE_T033_006') %>%
  dplyr::rename('% Not in Labor Force' = 'PCT_SE_T033_007') %>%
  ###Unemployment data(T37) left out since it is included in T33
   dplyr::rename('Median Household Income (In 2016 Inflation Adjusted Dollars)' = 'SE_T057_001') %>%
   dplyr::rename('Per Capita Income (In 2016 Inflation Adjusted Dollars)' = 'SE_T083_001') %>%
   dplyr::rename('Gini Index' = 'SE_T157_001') %>%
   dplyr::rename('Housing Units' = 'SE_T093_001') %>%
   dplyr::rename('Occupied Housing Units' = 'SE_T099_001') %>%
  ###unsure if rest of T99(housing fuel stuff) is useful 
  dplyr::rename('Median Gross Rent' = 'SE_T104_001') %>%
  dplyr::rename('Average Gross Rent for Renter-Occupied Housing Units' = 'SE_T106_001') %>%
  ###avg and median are actually somewhat different
  dplyr::rename('Population Under 18 Years of Age for Whom Poverty Status Is Determined' = 'SE_T114_001') %>%
  dplyr::rename('% Living in Poverty <18 determined' = 'PCT_SE_T114_002') %>%
  dplyr::rename('% At or Above Poverty Level <18 determined' = 'PCT_SE_T114_003') %>%
  dplyr::rename('Population Age 18 to 64 for Whom Poverty Status Is Determined' = 'SE_T115_001') %>%
  dplyr::rename('% Living in Poverty of 18-64 Determined' = 'PCT_SE_T115_002') %>%
  dplyr::rename('% At or Above Poverty Level of 18-64 Determined' = 'PCT_SE_T115_003') %>%
  dplyr::rename('Population Age 65 and Over for Whom Poverty Status Is Determined' = 'SE_T116_001') %>%
  dplyr::rename('% Living in Poverty 65+ Determined' = 'PCT_SE_T116_002') %>%
  dplyr::rename('% At or Above Poverty Level 65+ Determined' = 'PCT_SE_T116_003') 

  
  ###add more descriptions for each of the percentage ones
### income poverty level variable hard to interpret
  
```

Oregon County Census Data - Grouping By Area

```{r}

perc_or_census_cleaned <- perc_or_census_cleaned %>%
  mutate(countyname2 = str_remove(Geo_NAME, pattern = ' County, Oregon'))
  

perc_or_census_cleaned$Area <- perc_or_census_cleaned$countyname2

perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Hood River')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Jefferson')] <- 'Central_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Crook')] <- 'Central_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Tillamook')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Curry')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Wasco')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Columbia')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Malheur')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Union')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Clatsop')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Lincoln')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Umatilla')] <- 'Rural_Oreogn'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Coos')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Polk')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Yamhill')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Klamath')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Josephine')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Douglas')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Linn')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Deschutes')] <- 'Central_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Benton')] <- 'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Jackson')] <- 'Jackson'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Clackamas')] <- 'Clackamas'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Marion')] <-'Rural_Oregon'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Washington')] <- 'Washington'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Lane')] <- 'Lane'
perc_or_census_cleaned$Area[which(perc_or_census_cleaned$countyname2 == 'Multnomah')] <- 'Multnomah' 
perc_or_census_cleaned <- perc_or_census_cleaned %>%
  filter(Area == 'Lane' | Area == 'Multnomah' | Area == 'Clackamas' | Area == 'Rural_Oregon' | Area == 'Central_Oregon' | Area == 'Washington' | Area == 'Jackson')



perc_or_census_grouped <- perc_or_census_cleaned %>%
  group_by(Area) %>%
  dplyr::summarise(`% Unemployed` = mean(`% Unemployed`), `Gini Index` = mean(`Gini Index`), `Housing Units` = sum(`Housing Units`), `Median Gross Rent` = mean(`Median Gross Rent`), `Highschool Dropout Rate` = mean(`% Not High School Graduate, Not Enrolled (Dropped Out)`), `Total Population` = sum(`Total Population`), `% In Poverty` = sum(`% Living in Poverty of 18-64 Determined`) + sum(`% Living in Poverty 65+ Determined`))

```


#All States Census
```{r}
setwd("~/counting-the-homeless/Data")

states_census <- read.csv('R11689860_SL040.csv')
states_census_cleaned <- states_census %>%
  select(1:4, 56:143) %>%
  dplyr::rename('Total Population' = 'SE_T001_001') %>%
  dplyr::rename('Civilian Population 16 to 19 Years' = 'SE_T030_001') %>%
  dplyr::rename('% Not High School Graduate, Not Enrolled (Dropped Out)' = 'PCT_SE_T030_002') %>%
  dplyr::rename('% High School Graduate, or Enrolled (in School)' = 'PCT_SE_T030_003') %>%
  dplyr::rename('Population 16 Years and Over' = 'SE_T033_001') %>% 
  dplyr::rename('% In Labor Force' = 'PCT_SE_T033_002') %>%
  dplyr::rename('% In Armed Forces' = 'PCT_SE_T033_003') %>%
  dplyr::rename('% Civilian' = 'PCT_SE_T033_004') %>%
  dplyr::rename('% Employed' = 'PCT_SE_T033_005') %>%
  dplyr::rename('% Unemployed' = 'PCT_SE_T033_006') %>%
  dplyr::rename('% Not in Labor Force' = 'PCT_SE_T033_007') %>%
  ###Unemployment data(T37) left out since it is included in T33
   dplyr::rename('Median Household Income (In 2016 Inflation Adjusted Dollars)' = 'SE_T057_001') %>%
   dplyr::rename('Per Capita Income (In 2016 Inflation Adjusted Dollars)' = 'SE_T083_001') %>%
   dplyr::rename('Gini Index' = 'SE_T157_001') %>%
   dplyr::rename('Housing Units' = 'SE_T093_001') %>%
   dplyr::rename('Occupied Housing Units' = 'SE_T099_001') %>%
  ###unsure if rest of T99(housing fuel stuff) is useful 
  dplyr::rename('Median Gross Rent' = 'SE_T104_001') %>%
  dplyr::rename('Average Gross Rent for Renter-Occupied Housing Units' = 'SE_T106_001') %>%
  ###avg and median are actually somewhat different
  dplyr::rename('Population Under 18 Years of Age for Whom Poverty Status Is Determined' = 'SE_T114_001') %>%
  dplyr::rename('% Living in Poverty <18 determined' = 'PCT_SE_T114_002') %>%
  dplyr::rename('% At or Above Poverty Level <18 determined' = 'PCT_SE_T114_003') %>%
  dplyr::rename('Population Age 18 to 64 for Whom Poverty Status Is Determined' = 'SE_T115_001') %>%
  dplyr::rename('% Living in Poverty of 18-64 Determined' = 'PCT_SE_T115_002') %>%
  dplyr::rename('% At or Above Poverty Level of 18-64 Determined' = 'PCT_SE_T115_003') %>%
  dplyr::rename('Population Age 65 and Over for Whom Poverty Status Is Determined' = 'SE_T116_001') %>%
  dplyr::rename('% Living in Poverty 65+ Determined' = 'PCT_SE_T116_002') %>%
  dplyr::rename('% At or Above Poverty Level 65+ Determined' = 'PCT_SE_T116_003') 
```


# Grouping the state census data
```{r}
states_census_grouped <- states_census_cleaned %>%
  group_by(Geo_NAME) %>%
  dplyr::summarise(`% Unemployed` = mean(`% Unemployed`), `Gini Index` = mean(`Gini Index`), `Housing Units` = sum(`Housing Units`), `Median Gross Rent` = mean(`Median Gross Rent`), `Highschool Dropout Rate` = mean(`% Not High School Graduate, Not Enrolled (Dropped Out)`), `Total Population` = sum(`Total Population`), `% In Poverty` = sum(`% Living in Poverty of 18-64 Determined`) + sum(`% Living in Poverty 65+ Determined`))


```


### Ultimate Datasets

```{r}

Ultimate_Oregon <- inner_join(perc_or_census_grouped, PIT_Housing_Oregon_Merge, by = c('Area' = 'Area')) 

Ultimate_Oregon$Area[which(Ultimate_Oregon$Area == 'Central_Oregon')] <- 'Central Oregon'
Ultimate_Oregon$Area[which(Ultimate_Oregon$Area == 'Rural_Oregon')] <- 'Rural Oregon'

Ultimate_States <- inner_join(PIT_Housing_State_Merge, states_census_grouped, by = c('State_Name' = 'Geo_NAME'))


```